---
title: "length"
output: html_document
---

We want to know the length of a shared segment of DNA that could be observed with no recombination with some probability. The probability that a sequence of length $L$ will have no recombination in one generation is $P(no rec over length L)=(1-c)^L$ where $c$ is the recombination rate per bp. Then, the probability that this locus will have no recombination for $n$ generations is simply $P(no rec over length L over n gens)=((1-c)^L)^n$.  If we set that probability to some value $P$ we can estimate the length of the segement for some value of $n$ via: $L=\frac{log(P)}{n*log(1-c)}$.

```{r}
rec<-function(power,gens,c){
 return(log(power)/(gens*log(1-c)))  
}
```

We know do this for a bunch of values of time for three values of the recombination rate, 1cM/Mb (1E-8 per bp or average-ish recombination), 0.1cM/Mb (low recombination), and 10cM/Mb (high recombination)
```{r}
power=0.8
c=1E-8 #rec per bp

length1=sapply(1:2000*100,function(x) rec(power,x,c/10))/1000
length2=sapply(1:2000*100,function(x) rec(power,x,c))/1000
length3=sapply(1:2000*100,function(x) rec(power,x,c*10))/1000
g=1:2000*100/1000 #generations
```

Plot over whole range.
```{r, echo=FALSE}
par(mar=c(5,5,3,3))
plot(length1~g,log="y",type="line",lty=1,lwd=2,xlab="Time (ky)",cex.lab=1.5,ylab="Length (kb)",ylim=c(0.1,2000),yaxt="n",cex.axis=1.25)
axis(side=2,at=c(.1,1,5,10,50,100,500),labels=c(.1,1,5,10,50,100,500),cex.axis=1.25)
legend("top",lty=c(1,2,3),legend=c(0.1,1,10),box.lwd=0,title="cM/Mb",lwd=2)
lines(length2~g,log="y",type="line",lty=2,lwd=2)
lines(length3~g,log="y",type="line",lty=3,lwd=2)
```

Plot over 1-5k year range.
```{r, echo=FALSE}
par(mar=c(5,5,3,3))
plot(length1~g,log="y",type="line",lty=1,lwd=3,ylab="Length (kb)",cex.lab=2,xlab="Time (ky)",ylim=c(0.1,2000),yaxt="n",cex.axis=1.25,xlim=c(0,5))
axis(side=2,at=c(.1,1,5,10,50,100,500),labels=c(.1,1,5,10,50,100,500),cex.axis=1.25)
lines(length2~g,log="y",type="line",lty=2,lwd=3)
lines(length3~g,log="y",type="line",lty=3,lwd=3)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
